---
title: "DS2 Midterm Project"
author: "Siying Li"
date: "3/31/2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(ggplot2)
library(dplyr)
library(caret)
library(gridExtra)
library(tidyverse)
library(glmnet)
library(corrplot)
library(splines)
library(lasso2)
library(Rmisc)
library(FNN)
library(pdp)
library(earth)
library(sandwich)
library(stargazer)

```

## 1 Data Visualization
```{r}
house = read.csv(file = "train.csv", stringsAsFactors = FALSE)
# Do not import strings as factors, since the ultimate goal is to transfer all variables to numeric. 

dim(house)

house = house %>% 
  dplyr::select(-Id) %>% 
  janitor::clean_names()
```

The house dataset consists of both integer and character variables. Most of the categorical variables are ordinal. There is a total of 81 variables, and the last column is our response. 

#### 1.1 Data Cleaning
```{r}
# Get rid of virables with 500 plus missing values
missing = 
  colSums(sapply(house, is.na)) %>% 
  as.data.frame() %>% 
  mutate(variable = colnames(house)) %>% 
  filter(. > 500) %>% 
  pull(variable)

house = 
  house %>% 
  select(-missing) %>% 
  filter(lot_area < 100000) # filter out house with extreme lot size to prevent outliers. 

```

#### 1.2 Response Variable
```{r}
# Response variable distribution
p1 = ggplot(data=house, aes(x=sale_price)) +
        geom_histogram(fill="pink", binwidth = 10000) +
        scale_x_continuous(breaks= seq(0, 800000, by=100000)) + ggtitle("Response Distribution")
  
# Right skewed

# log transform on y
p2 = ggplot(data=house, aes(x=log(sale_price))) +
        geom_histogram(fill="pink", binwidth = 0.05) +
        scale_x_continuous(breaks= seq(0, 800000, by=100000))+ ggtitle("log(Response) Distribution")

grid.arrange(p1, p2, ncol = 2)

```


#### 1.3 Numeric Predictors
```{r}
label_num <- sapply(house, is.numeric) #set numeric variables as TRUE
num_var <- house[, label_num] # get all the numeric variables 
```

There is a total of `r length(num_var)` numeric variables. 

```{r}
# Test Correlation
corr_num <- cor(num_var, use = "pairwise.complete.obs") #correlations of all numeric variables

#sort on decreasing correlations with sale price
corr_sort <- as.matrix(sort(corr_num[,'sale_price'], decreasing = TRUE))
#eliminate low correlation variables 
high_corr <- names(which(apply(corr_sort, 1, function(x) abs(x)>0.3)))
corr_num <- corr_num[high_corr, high_corr]

corrplot.mixed(corr_num, tl.col="black", tl.pos = "lt")

# 'Lot_frontage' variable represent linear feet of street connected to property. It contains NAs, which represents there is no backyard. Replace NA with zero. 
house$lot_frontage[is.na(house$lot_frontage)] = 0

# There is a "1st floor sq footage" variable and a "2nd floor square footage" variable.
# It would be better to sum them up to total square footage for better pridiction. 
house = 
  house %>% 
  mutate(total_sq = x1st_flr_sf + x2nd_flr_sf) %>% 
  select(-x1st_flr_sf, -x2nd_flr_sf)

# Convert year and month sold to factor variable
house$mo_sold = as.factor(house$mo_sold)
house$yr_sold = as.factor(house$yr_sold)

```


#### 1.4 Categorical Predictors
```{r}
char_num <- names(house[,sapply(house, is.character)])
length(char_num)


# Oridinal Predictors 
house[house == "Ex"] = 5
house[house == "Gd"] = 4
house[house == "TA"] = 3
house[house == "Fa"] = 2
house[house == "Po"] = 1
house[house == "NA"] = 3

house$exter_qual = as.numeric(house$exter_qual)
house$exter_qual = as.numeric(house$exter_qual)
house$bsmt_qual = as.numeric(house$bsmt_qual)
house$bsmt_cond = as.numeric(house$bsmt_cond)
house$heating_qc= as.numeric(house$heating_qc)
house$kitchen_qual= as.numeric(house$kitchen_qual)
house$garage_qual= as.numeric(house$garage_qual)
house$garage_cond= as.numeric(house$garage_cond)


# LotShape: General shape of property
house$lot_shape[house$lot_shape == "Reg"] = 4 # Regular
house$lot_shape[house$lot_shape == "IR1"] = 3 # Slightly irregular
house$lot_shape[house$lot_shape == "IR2"] = 2 # Moderately Irregular
house$lot_shape[house$lot_shape == "IR3"] = 1 # Irregular
house$lot_shape = as.numeric(house$lot_shape)

# Utilities: Type of utilities available
house$utilities[house$utilities == "AllPub"] = 4 # All public Utilities (E,G,W,& S)
house$utilities[house$utilities == "NoSewr"] = 3 # Electricity, Gas, and Water (Septic Tank)
house$utilities[house$utilities == "NoSeWa"] = 2 # Electricity and Gas Only
house$utilities[house$utilities == "ELO"] = 1    # Electricity only
house$utilities = as.numeric(house$utilities)

# LandSlope: Slope of property
house$land_slope[house$land_slope == "Gtl"] = 3 # Gentle slope
house$land_slope[house$land_slope == "Mod"] = 2 # Moderate Slope
house$land_slope[house$land_slope == "Sev"] = 1 # Severe Slope
house$land_slope = as.numeric(house$land_slope)

# BsmtExposure: Refers to walkout or garden level walls
house$bsmt_exposure[house$bsmt_exposure == "Av"] = 3 # Average Exposure (split levels or foyers typically score average or above)  
house$bsmt_exposure[house$bsmt_exposure == "Mn"
                    ] = 2 # Mimimum Exposure
house$bsmt_exposure[house$bsmt_exposure == "No"] = 1 # No Exposure
house$bsmt_exposure[house$bsmt_exposure == 0] = 0 # No Exposure
house$bsmt_exposure = as.numeric(house$bsmt_exposure)

# BsmtFinType1: Rating of basement finished area
house$bsmt_fin_type1[house$bsmt_fin_type1 == "GLQ"] = 6 # Good Living Quarters
house$bsmt_fin_type1[house$bsmt_fin_type1 == "ALQ"] = 5 # Average Living Quarters
house$bsmt_fin_type1[house$bsmt_fin_type1 == "BLQ"] = 4 # Below Average Living Quarters
house$bsmt_fin_type1[house$bsmt_fin_type1 == "Rec"] = 3 # Average Rec Room
house$bsmt_fin_type1[house$bsmt_fin_type1 == "LwQ"] = 2 # Low Quality
house$bsmt_fin_type1[house$bsmt_fin_type1 == "Unf"] = 1 # Unfinshed
house$bsmt_fin_type1[house$bsmt_fin_type1 == "NA"] = 3 # No Basement
house$bsmt_fin_type1 = as.numeric(house$bsmt_fin_type1)

# CentralAir: Central air conditioning
house$central_air[house$central_air == "N"] = 0 # No
house$central_air[house$central_air == "Y"] = 1 # Yes
house$central_air = as.numeric(house$central_air)

# Functional: Home functionality (Assume typical unless deductions are warranted)
house$functional[house$functional == "Typ"] = 8 # Typical Functionality
house$functional[house$functional == "Min1"] = 7 # Minor Deductions 1
house$functional[house$functional == "Min2"] = 6 # Minor Deductions 2
house$functional[house$functional == "Mod"] = 5 # Moderate Deductions
house$functional[house$functional == "Maj1"] = 4 # Major Deductions 1
house$functional[house$functional == "Maj2"] = 3 # Major Deductions 2
house$functional[house$functional == "Sev"] = 2 # Severely Damaged
house$functional[house$functional == "Sal"] = 1 # Salvage only
house$functional = as.numeric(house$functional)

# GarageFinish: Interior finish of the garage
house$garage_finish[house$garage_finish == "Fin"] = 3 # Finished
house$garage_finish[house$garage_finish == "RFn"] = 2 # Rough Finished
house$garage_finish[house$garage_finish == "Unf"] = 1 # Unfinished
house$garage_finish[house$garage_finish == 3 ] = 0 # No Garage
house$garage_finish = as.numeric(house$garage_finish)

# PavedDrive: Paved driveway
house$paved_drive[house$paved_drive == "Y"] = 3 # Paved 
house$paved_drive[house$paved_drive == "P"] = 2 # Partial Pavement
house$paved_drive[house$paved_drive == "N"] = 1 # Dirt/Gravel
house$paved_drive = as.numeric(house$paved_drive)

# sapply(house, class)

# Select rest of the character variables and change them to factors. 
house[sapply(house, is.character)] <- lapply(house[sapply(house, is.character)], 
                                       as.factor)
# Select rest of the integer variables and change them to numeric. 
# house[sapply(house, is.integer)] <- lapply(house[sapply(house, is.integer)], 
#                                        as.numeric)
```


#### 1.5 Remove near zero variance pridictors and rows contaning NA
```{r}
# There are many variables that contains many zeros. Use a 95% cutoff for the percentage of distinct values out of the number of total samples. 
near_zero = 
  house %>% 
  nearZeroVar(names = TRUE, freqCut = 75/25)

house = 
  house %>% 
  select(-near_zero) %>% 
  drop_na()

```

Finally, numer of numeric variable is `r length(house[,sapply(house, is.numeric)])`. And the numer of factor variable is `r length(house[,sapply(house, is.factor)])`.

## 2 Model Fitting
```{r}
x = model.matrix(sale_price~., house) [,-1]
y = log(house$sale_price)

# remove colinear
linear_combo = findLinearCombos(x)
x = x[, -linear_combo$remove]

# remove near zero variance
near_zero_x = x %>% nearZeroVar(names = TRUE, freqCut = 85/15)
x = as.data.frame(x)
x = x %>% select(-near_zero_x)
x = data.matrix(x)

ctrl1 = trainControl(method = "repeatedcv", repeats = 5)
```

```{r}
theme1 <- trellis.par.get()
theme1$plot.symbol$col <- rgb(.2, .4, .2, .5)
theme1$plot.symbol$pch <- 16
theme1$plot.line$col <- rgb(.8, .1, .1, 1)
theme1$plot.line$lwd <- 2
theme1$strip.background$col <- rgb(.0, .2, .6, .2)
trellis.par.set(theme1)
featurePlot(x, y, plot = "scatter", labels = c("","Y"),
            type = c("p"), layout = c(7, 7))
```


### 2.1 Multiple Linear Regression
```{r}
set.seed(2)

lm.fit = train(x, y, method = "lm", trControl = ctrl1, preProcess = c("center", "scale"))

# MSE
lm.fit$results$RMSE
```

### 2.2 Ridge Regression
```{r}
set.seed(2)

ridge.fit <- train(x, y,
                   method = "glmnet",
                   tuneGrid = expand.grid(alpha = 0, 
                                          lambda = exp(seq(-4, -2, length=200))),
                   preProc = c("center", "scale"),
                   trControl = ctrl1)

plot(ridge.fit, xTrans = function(x) log(x))

# ridge.fit$results$RMSE

ridge.fit$bestTune

# coef(ridge.fit$finalModel,ridge.fit$bestTune$lambda)
```

### 2.3 Lasso
```{r}
set.seed(2)

lasso.fit <- train(x, y,
                   method = "glmnet",
                   tuneGrid = expand.grid(alpha = 1, 
                                          lambda = exp(seq(-9,-3, length=200))),
                   preProc = c("center", "scale"),
                   trControl = ctrl1)

plot(lasso.fit, xTrans = function(x) log(x))

lasso.fit$bestTune

# number of non-zero coefficient
coef = coef(lasso.fit$finalModel,lasso.fit$bestTune$lambda)
nnzero(coef)

# coef(lasso.fit$finalModel,lasso.fit$bestTune$lambda)
```

### 2.4 Elastic Net
```{r}
set.seed(2)

enet.fit <- train(x, y,
                  method = "glmnet",
                  tuneGrid = expand.grid(alpha = seq(0, 1, length = 5), 
                                         lambda = exp(seq(-9, -2, length = 50))),
                  preProc = c("center", "scale"),
                  trControl = ctrl1)
enet.fit$bestTune

ggplot(enet.fit, highlight = TRUE)
```

### 2.5 Principal Component Repression (PCR)
```{r}
set.seed(2)

pcr.fit <- train(x, y,
                 method = "pcr",
                 tuneGrid  = data.frame(ncomp = 1:ncol(x)),
                 trControl = ctrl1,
                 preProc = c("center", "scale"))

ggplot(pcr.fit, highlight = TRUE) + theme_bw()
```

### 2.6 Partial Least Squares
```{r}
set.seed(2)
pls.fit <- train(x, y,
                 method = "pls",
                 tuneGrid = data.frame(ncomp = 1:ncol(x)),
                 trControl = ctrl1,
                 preProc = c("center", "scale"))

ggplot(pls.fit, highlight = TRUE)
```

### 2.7 GAM
```{r}
set.seed(2)
gam.fit <- train(x, y,
                 method = "gam",
                 tuneGrid = data.frame(method = "GCV.Cp", select = c(TRUE,FALSE)),
                 trControl = ctrl1)
gam.fit$finalModel

```

### 2.8 Multivariable Adaptive Regression Splines (MARS)
```{r}
set.seed(2)
mars_grid <- expand.grid(degree = 1:2, 
                         nprune = 2:ncol(x))

set.seed(2)
mars.fit <- train(x, y,
                  method = "earth",
                  tuneGrid = mars_grid,
                  trControl = ctrl1)

ggplot(mars.fit)

mars.fit$bestTune

coef(mars.fit$finalModel) 


plot_mars = function(variable)
  { p1 <- partial(mars.fit, pred.var = c(variable), grid.resolution = 10) %>% autoplot()
  p1}

name = c("overall_qual", "total_sq", "year_built", "fireplaces", "year_remod_add", "bsmt_fin_sf1", "open_porch_sf", "lot_area", "total_bsmt_sf")

plot_list = map(name,plot_mars)

multiplot(plotlist = plot_list, cols=3)
```

## 3 Between Model Comparison
```{r}
resamp <- resamples(list(Elastic = enet.fit, Lasso = lasso.fit, Ridge = ridge.fit, LM = lm.fit, PCR = pcr.fit, PLS = pls.fit, GAM = gam.fit, MARS = mars.fit))
summary(resamp)

parallelplot(resamp, metric = "RMSE")
bwplot(resamp, metric = "RMSE")
```

## 4 Final Model Exploration
```{r}
varImp(mars.fit)
```


